\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-01-06 manuscript.Rnw
Minor text edits

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and QALYs from a PSA.
}

@@

Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}). The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}. It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}. Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to cohort Markov models.

>>>>

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and utility in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years (QALYs) from a PSA.
}

@@

Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}). The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}. It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}. Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited Markov cohort models.
